---
author:
- Sarah Gaichas, PhD
- Research Fishery Biologist
- Ecosystem Dynamics and Assessment Branch
- Northeast Fisheries Science Center
opening: To the SSC,
closing: Sincerely,
date: "`r format(Sys.time(), '%d %B, %Y')`"
address:
- Mid-Atlantic Fishery Management Council
- Scientific and Statistical Committee

letterhead: letterhead/NEFSCletterhead.pdf

output: linl::linl
---

## Introduction

In this memo we list the comments and requests received on the 2019 State of the Ecosystem (SOE) reports, and how we responded to those requests. We include comments from both Councils because adjustments to the report were made in response to both. We welcome comments on whether this memo is useful and how to imporve it for future SOE reporting.

## Comment summary table

In the table below we summarize all comments and requests with sources. The Progress column briefly summarizes how we responded, with a more detailed response in the numbered Memo Section. In the Progress column, "SOE" indicates a change included in the report(s).

```{r summtable}
# take from google drive to automatically update any changes
googledrive::drive_download("Northeast IEA/State of the Ecosystem Reports/SOE 2020/Workshop & Meeting Notes/Request checklist", path="2019requests.xlsx", overwrite = TRUE)

knitr::kable(read_excel("2019request.xlsx", sheet="MemoTable", range = 'A1:D32')) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```


## Responses to comments



```{r, echo=FALSE}
#remotes::install_github("gadenbuie/lorem")
```

`r lorem::ipsum(paragraphs = 2)`
