---
title:
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
output: 
  pdf_document:
    includes:
      in_header: latex/header2021.tex
    keep_tex: yes
bibliography: SOE2020.bib
csl: plos.csl
link-citations: yes
fontsize: 10pt
subparagraph: yes
---

```{r setup, include=FALSE}
#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      dev = "cairo_pdf",
                      fig.align = 'center') #allows for inserting R code into captions

library(tidyverse)
library(readxl)
library(kableExtra)
library(ecodata)
library(googledrive)
library(pacman)
p_load(repmis, RCurl, DT, magrittr, gsubfn, stringr, tidyverse)


```

```{r, include = F}
#this is to wrap text around figures if we want
defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

NOTES to be REMOVED

Read in the google sheet with the list of requests

```{r, message=FALSE}
#results link https://docs.google.com/spreadsheets/d/1BSsibzyrj-ge3n55BunQBgNPbcL_gy3BzwGmosnV2vY/edit#gid=0

resultfile <- drive_find(pattern = "SOE_Request_Checklist", type = "spreadsheet")

requests <- drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
  {read.csv(.$local_path)} 

```


Use the code=readLines with the setup chunks from ecodata: need to do before each set of figs from a region/subgroup

# Introduction

In the table below we summarize all comments and requests with sources. The Progress column briefly summarizes how we responded, with a more detailed response in the numbered Memo Section. In the Progress column, "SOE" indicates a change included in the report(s).

```{r summtable}
# take from google drive to automatically update any changes
#googledrive::drive_download("Northeast IEA/State of the Ecosystem Reports/SOE 2020/Workshop & Meeting Notes/Request checklist", path="2019requests.xlsx", overwrite = TRUE)

#requests <- read_excel("2019requests.xlsx", sheet="MemoTable", range = 'A1:D31')

#requests$`Memo Section` <- paste0("[", requests$`Memo Section`, "](#",requests$`Memo Section`,")") #trying to make clickable links in the table to sections, not working
#adding format = "markdown" to kable statement made them links but they still didnt jump to report sections and messed up other formatting

requeststab <- requests %>%
  select(Request, Year, Source, Progress) %>%
  mutate("Memo Section" = rownames(.))

knitr::kable(requeststab, longtable = T, booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size=9, latex_options = c("hold_position", "striped", "repeat_header")) %>%
  row_spec(0,bold=TRUE) %>%
  column_spec(1, width="5cm") %>%
  column_spec(2, width="1cm") %>%
  column_spec(3, width="2cm") %>%
  column_spec(4, width="5cm") %>%
  column_spec(5, width="2cm")
```


# Responses to comments



# References

