---
title:
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
output: 
  pdf_document:
    includes:
      in_header: latex/header2021.tex
    keep_tex: yes
bibliography: SOE2021.bib
csl: plos.csl
link-citations: yes
fontsize: 10pt
subparagraph: yes
---

```{r setup, include=FALSE}
#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      dev = "cairo_pdf",
                      fig.align = 'center') #allows for inserting R code into captions

library(tidyverse)
library(readxl)
library(kableExtra)
library(ecodata)
library(googledrive)
library(pacman)
p_load(repmis, RCurl, DT, magrittr, gsubfn, stringr, tidyverse)


```

```{r, include = F}
#this is to wrap text around figures if we want
defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

<!--
NOTES

Use the code=readLines with the setup chunks from ecodata: need to do before each set of figs from a region/subgroup

-->

```{r, message=FALSE, results='hide'}
#results link https://docs.google.com/spreadsheets/d/1BSsibzyrj-ge3n55BunQBgNPbcL_gy3BzwGmosnV2vY/edit#gid=0

resultfile <- drive_find(pattern = "SOE_Request_Checklist", type = "spreadsheet")

requests <- drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
  {read.csv(.$local_path)} 

```

# Introduction

In the table below we summarize all comments and requests with sources. The Progress column briefly summarizes how we responded, with a more detailed response in the numbered Memo Section. In the Progress column, "SOE" indicates a change included in the report(s).

```{r summtable}
# take from google drive to automatically update any changes
#googledrive::drive_download("Northeast IEA/State of the Ecosystem Reports/SOE 2020/Workshop & Meeting Notes/Request checklist", path="2019requests.xlsx", overwrite = TRUE)

#requests <- read_excel("2019requests.xlsx", sheet="MemoTable", range = 'A1:D31')

#requests$`Memo Section` <- paste0("[", requests$`Memo Section`, "](#",requests$`Memo Section`,")") #trying to make clickable links in the table to sections, not working
#adding format = "markdown" to kable statement made them links but they still didnt jump to report sections and messed up other formatting

requeststab <- requests %>%
  select(Request, Year, Source, Progress) %>%
  mutate("Memo Section" = rownames(.))

knitr::kable(requeststab, longtable = T, booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size=9, latex_options = c("hold_position", "striped", "repeat_header")) %>%
  row_spec(0,bold=TRUE) %>%
  column_spec(1, width="5cm") %>%
  column_spec(2, width="1cm") %>%
  column_spec(3, width="2cm") %>%
  column_spec(4, width="5cm") %>%
  column_spec(5, width="2cm")
```

# Responses to comments

<!-- compiled this to md then copied back in, filled in text between
```{r, results='asis'}
for(i in 1:nrow(requeststab)) {
  cat("##", requeststab$'Memo Section'[i], as.character(requeststab$Request[i]), "\n")
  cat("\n")
}
```
-->

1 Report card and summary visualizations
----------------------------------------

Both Councils requested a "report card" style summary section with visualizations in 2019. We introduced a 2 page summary format in 2020 with a bulleted list of results on the first page and visualizations on the second. This year, the report was reorganized to more clearly link indicators with fishery management objectives and to better synthesize results across indicators, so the summary section was restructured accordingly. The 2021 summary pages include: 

1. a report card style table summarizing status and trends of indicators linked to management objectives, combined with brief descriptions of implications for management synthesizing across multiple indicators in the report;  
2. a bulleted list highlighting risks to meeting fishery management objectives, including those from climate change and those from wind energy development; and  
3. visualizations of ecosystem synthesis themes integrated in the report, including multiple drivers of change, regime shifts, and ecosystem reorganization.  


2 Ecosystem Overfishing indicators (Link and Watson, 2019)
----------------------------------------------------------

Both Councils have requested more information on ecosystem thresholds and inflection points. This year we have included proposed ecosystem overfishing indicators and thresholds [@link_global_2019]. 

3 Primary production required, interpretation of decline?
---------------------------------------------------------

4 Climate Change context
------------------------

5 Clarify language (e.g., primary production required)
------------------------------------------------------

6 Copy Editing
--------------

7 Ocean Acidification
---------------------

8 Include examples of High/Low engaged ports
--------------------------------------------

9 Expand wind lease area and habitat overlap
--------------------------------------------

10 Expand cold pool index
-------------------------

11 Seperate Bigelow/Albatross catch diversity metric
----------------------------------------------------

12 Shark diversity indicator
----------------------------

13 Uncertainty estimates
------------------------

14 Bycatch index
----------------

15 Marine Mammal consumption
----------------------------

16 Estuarine Water Quality
--------------------------

17 Linking Condition
--------------------

18 Avg weight of diet components by feeding group
-------------------------------------------------

19 Mean stomach weight across feeding guilds
--------------------------------------------

20 Shellfish growth/distribution linked to climate (system productivity)
------------------------------------------------------------------------

21 Cumulative weather index
---------------------------

22 Management complexity
------------------------

23 VAST and uncertainty
-----------------------

24 Seal index
-------------

25 Incorporate social sciences survey from council
--------------------------------------------------

26 Young of Year index from multiple surveys
--------------------------------------------

27 Biomass of spp not included in BTS
-------------------------------------

28 Estuarine condition relative to power plants and temp
--------------------------------------------------------

29 Inflection points for indicators
-----------------------------------

30 Reduce indicator dimensionality with multivariate statistics
---------------------------------------------------------------

31 Breakpoints
--------------

32 Re-evaluate EPUs
-------------------
# References

