---
title:
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
output: 
  pdf_document:
    includes:
      in_header: latex/header2021.tex
    keep_tex: yes
bibliography: SOE2021.bib
csl: plos.csl
link-citations: yes
fontsize: 10pt
subparagraph: yes
---

```{r setup, include=FALSE}
#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      dev = "cairo_pdf",
                      fig.align = 'center') #allows for inserting R code into captions

library(tidyverse)
library(readxl)
library(kableExtra)
library(ecodata)
library(googledrive)
library(pacman)
p_load(repmis, RCurl, DT, magrittr, gsubfn, stringr, tidyverse)


```

```{r, include = F}
#this is to wrap text around figures if we want
defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

<!--
NOTES

Use the code=readLines with the setup chunks from ecodata: need to do before each set of figs from a region/subgroup

-->

```{r, message=FALSE, results='hide'}
#results link https://docs.google.com/spreadsheets/d/1BSsibzyrj-ge3n55BunQBgNPbcL_gy3BzwGmosnV2vY/edit#gid=0

resultfile <- drive_find(pattern = "SOE_Request_Checklist", type = "spreadsheet")

requests <- drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
  {read.csv(.$local_path)} 

```

# Introduction

In the table below we summarize all comments and requests with sources. The Progress column briefly summarizes how we responded, with a more detailed response in the numbered Memo Section. In the Progress column, "SOE" indicates a change included in the report(s).

```{r summtable}
# take from google drive to automatically update any changes
#googledrive::drive_download("Northeast IEA/State of the Ecosystem Reports/SOE 2020/Workshop & Meeting Notes/Request checklist", path="2019requests.xlsx", overwrite = TRUE)

#requests <- read_excel("2019requests.xlsx", sheet="MemoTable", range = 'A1:D31')

#requests$`Memo Section` <- paste0("[", requests$`Memo Section`, "](#",requests$`Memo Section`,")") #trying to make clickable links in the table to sections, not working
#adding format = "markdown" to kable statement made them links but they still didnt jump to report sections and messed up other formatting

requeststab <- requests %>%
  select(Request, Year, Source, Progress) %>%
  mutate("Memo Section" = rownames(.))

knitr::kable(requeststab, longtable = T, booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size=9, latex_options = c("hold_position", "striped", "repeat_header")) %>%
  row_spec(0,bold=TRUE) %>%
  column_spec(1, width="5cm") %>%
  column_spec(2, width="1cm") %>%
  column_spec(3, width="2cm") %>%
  column_spec(4, width="5cm") %>%
  column_spec(5, width="2cm")
```

# Responses to comments

<!-- compiled this to md then copied back in, filled in text between
```{r, results='asis'}
for(i in 1:nrow(requeststab)) {
  cat("##", requeststab$'Memo Section'[i], as.character(requeststab$Request[i]), "\n")
  cat("\n")
}
```
-->

1 Report card and summary visualizations
----------------------------------------

Both Councils requested a "report card" style summary section with visualizations in 2019. We introduced a 2 page summary format in 2020 with a bulleted list of results on the first page and visualizations on the second. This year, the report was reorganized to more clearly link indicators with fishery management objectives and to better synthesize results across indicators, so the summary section was restructured accordingly. The 2021 summary pages include: 

1. a report card style table summarizing status and trends of indicators linked to management objectives, combined with brief descriptions of implications for management synthesizing across multiple indicators in the report;  
2. a bulleted list highlighting risks to meeting fishery management objectives, including those from climate change and those from wind energy development; and  
3. visualizations of ecosystem synthesis themes integrated in the report, including multiple drivers of change, regime shifts, and ecosystem reorganization.  

We welcome feedback on these revisions and suggestions for further refinements to make this summary more useful.

2 Ecosystem Overfishing indicators (Link and Watson, 2019)
----------------------------------------------------------

Both Councils have requested more information on ecosystem thresholds and inflection points. This year we have calculated two ecosystem overfishing indicators with proposed thresholds [@link_global_2019] for each ecological production unit (EPU) on the northeast US shelf. 

3 Primary production required, interpretation of decline?
---------------------------------------------------------



4 Climate Change context
------------------------

The NE SSC was interested in more explicitly addressing climate change in the reports. As described above, we have now reorganized the report into two major sections. The second section outlines risk to meeting fishery management objectives, with climate change representing the first major risk category (the other is offshore wind energy development). Climate risks to meeting fishery management objectives are also explicitly indicated in the first section on performance against management objectives. We welcome feedback on this structural revision.

Cimate forecasts at scales relevant to fishery management (months to years) are in progress, with at least one paper in review at present. We plan to include more of this information in future reports. 

5 Clarify language (e.g., primary production required)
------------------------------------------------------

Both Councils asked for clarification of several new terms, including "primary production required," and the NE SSC suggested adding a glossary to improve clarity. We have added an online glossary (https://noaa-edab.github.io/tech-doc/glossary.html) which is linked from the report to explain many terms. The Northeast Fisheries Science Center Research Communications Branch (NEFSC RCB) also reviewed the draft document to streamline language.

6 Copy Editing
--------------



7 Ocean Acidification
---------------------



8 Include examples of High/Low engaged ports
--------------------------------------------



9 Expand wind lease area and habitat overlap
--------------------------------------------



10 Expand cold pool index
-------------------------



11 Seperate Bigelow/Albatross catch diversity metric
----------------------------------------------------

The NE SSC requested a species diversity metric based on NEFSC trawl survey data. We had included such a metric in past reports (2017), but were concerned that apparent differences in diversity prior to and after 2008 may be driven by differences in survey vessels. While species-specific cpue and sizes have calibration coefficents between survey vessels, the number of species captured by the vessels has no known calibration coefficient.

After discussion with both SSCs in 2020, we calculated NEFSC trawl survey diversity metrics separately for the Albatross and Bigelow survey vessel time series. In each 2021 SOE we report the expected number of species per 1000 individuals sampled for each EPU in the fall, with uncertainty  (p.  MAFMC and p.  NEFMC). Distinguising potential vessel effects from trends in diversity should be facilitated by this presentation. Plots for spring, as well as comparisons with Shannon diversity metrics combining both vessel time series as originally calculated, are available online (https://noaa-edab.github.io/ecodata/macrofauna_NE#Survey_Shannon_Diversity, https://noaa-edab.github.io/ecodata/macrofauna_MAB#Survey_Shannon_Diversity). We welcome further discussion to refine this and other diversity indices.

12 Shark abundance and catch indicators
---------------------------------------

The MAFMC requested information on biomass of sharks, as fishermen had reported encountering more blacktip, spinner, and sandbar sharks each summer. We were able to obtain catch and CPUE data from the Highly Migratory Species group at NMFS Headquarters as well as bycatch information from the NEFSC Observer Program. 

13 Uncertainty estimates
------------------------

14 Bycatch index
----------------

15 Marine Mammal consumption
----------------------------

16 Estuarine Water Quality
--------------------------

17 Linking Condition
--------------------

18 Avg weight of diet components by feeding group
-------------------------------------------------

19 Mean stomach weight across feeding guilds
--------------------------------------------

20 Shellfish growth/distribution linked to climate (system productivity)
------------------------------------------------------------------------

21 Cumulative weather index
---------------------------

22 Management complexity
------------------------

The MAFMC asked for indicators of management complexity for use in the EAFM risk assessment. An NEFSC summer student started work on this in 2018, but we have lacked capacity to finish the project since then. If resources allow we will continue the project, and guidance for further indicator developmet is welcome.

23 VAST and uncertainty
-----------------------

24 Seal index
-------------

25 Incorporate social sciences survey from council
--------------------------------------------------

The NE SSC was interested in reviewing information on the perception and use of social science information from an NEFMC survey. We had insufficient resources to address this in 2020.  We welcome input from the New England Council and staff on how best to incorporate this information in future reports.

26 Young of Year index from multiple surveys
--------------------------------------------

The MA SSC was interested in a young of year index from multiple surveys. In past reports we have included the fish productivity index, which calculates the number of small fish per biomass of large fish of the same species from NEFSC surveys. This index is based only on the NEFSC bottom trawl survey, which was cancelled in 2020, so the index was not updated. We recognize that this is not strictly a young of year index, and it is from a single survey. 

We had insufficient resources to address this in 2020.  

27 Biomass of species not included in bottom trawl surveys
----------------------------------------------------------

We included information on sharks this year, and data streams for many other species not captured by bottom trawl surveys (BTS) are under investigation. However, we had insufficient resources to address this fully in 2020.  

28 Estuarine condition relative to power plants and temp
--------------------------------------------------------

We had insufficient resources to address this in 2020.  

29 Inflection points for indicators
-----------------------------------

While this could not be addressed for individual indicators in 2020, we did include new Ecosystem Overfishing indicators with proposed thresholds (see Section 2, [Ecosystem Overfishing indicators](#ecosystem-overfishing-indicators-link-and-watson-2019)). We welcome suggestions for which additional indicators or groups of indicators should be prioritized for inflection point/threshold analysis in upcoming years.  

30 Reduce indicator dimensionality with multivariate statistics
---------------------------------------------------------------

We had insufficient resources to address this in 2020.  

31 Breakpoints
--------------

While this could not be addressed for individual indicators in 2020, our newly introduced regime shifts synthesis theme will be explored further in upcoming years. We welcome suggestions for which individual indicators or groups of indicators should be prioritized for regime shift analysis in upcoming years.  

32 Re-evaluate EPUs
-------------------

Initial planning for re-evaluating Northeast US Shelf ecological production units has started, but we had insufficient resources to begin the project in 2020.  

# References

