---
title:
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
output: md_document
bibliography: SOE2020.bib
csl: plos.csl
link-citations: yes
fontsize: 10pt
subparagraph: yes
always_allow_html: true
---

```{r setup, include=FALSE}
#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      dev = "cairo_pdf",
                      fig.align = 'center') #allows for inserting R code into captions

library(tidyverse)
library(readxl)
library(kableExtra)
library(ecodata)
library(googledrive)
library(pacman)
p_load(repmis, RCurl, DT, magrittr, gsubfn, stringr, tidyverse)


```

```{r, include = F}
#this is to wrap text around figures if we want
defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

<!--
NOTES

Use the code=readLines with the setup chunks from ecodata: need to do before each set of figs from a region/subgroup
-->

```{r, results='hide'}
  resultfile <- googledrive::drive_find(pattern = "SOE 2021 Response Checklist", type = "spreadsheet")
 # 
  requests <- googledrive::drive_download(resultfile, type = "csv", overwrite = TRUE) %>%
    {read.csv(.$local_path)} 

#requests <- read.csv(here("SOE 2021 Response Checklist.csv"))
```
# Introduction

In the table below we summarize all comments and requests with sources. The Progress column briefly summarizes how we responded, with a more detailed response in the numbered Memo Section. In the Progress column, "SOE" indicates a change included in the report(s).

```{r summtable}
# take from google drive to automatically update any changes
#googledrive::drive_download("Northeast IEA/State of the Ecosystem Reports/SOE 2020/Workshop & Meeting Notes/Request checklist", path="2019requests.xlsx", overwrite = TRUE)

#requests <- read_excel("2019requests.xlsx", sheet="MemoTable", range = 'A1:D31')

#requests$`Memo Section` <- paste0("[", requests$`Memo Section`, "](#",requests$`Memo Section`,")") #trying to make clickable links in the table to sections, not working
#adding format = "markdown" to kable statement made them links but they still didnt jump to report sections and messed up other formatting

requeststab <- requests %>%
  select(Request, Year, Source, Status, Progress) %>%
  mutate(Source = str_trim(str_extract(Source, "[^-]+"))) %>%
  arrange(match(Status, c("In SOE", 
                          "In SOE-MAFMC",
                          "In SOE-MAFMC, In progress-NEFMC",
                          "In progress", 
                          "Not started")), desc(Year)) %>%
  mutate("Memo Section" = rownames(.))

knitr::kable(requeststab, longtable = T, booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size=9, latex_options = c("hold_position", "striped", "repeat_header")) %>%
  row_spec(0,bold=TRUE) %>%
  column_spec(1, width="5cm") %>%
  column_spec(2, width="1cm") %>%
  column_spec(3, width="2cm") %>%
  column_spec(4, width="2cm") %>%
  column_spec(5, width="5cm") %>%
  column_spec(6, width="2cm")
```


# Responses to comments

## 1 Add “This report is direct to…”

The first sentence of each report now states the audience.

## 2 State management objectives first in report

The graphical summary on page 2 (first content page for MAFMC) and 2-3 (first content pages for NEFMC) summarize management objectives (columns in each table) and the indicator's performance relative to those objectives. Table 1 in the main body each report lists management objectives. 

## 3 OA in NEFMC SOE

## 4 Habitat impact of fishing based on gear.

## 5 Revisit right whale language

## 6 Fishery gear modifications accounted for in shark CPUE? ask HMS

## 7 Sum of TAC/ Landings relative to TAC

## 8 Estuarine Water Quality

## 9 More direct convos for feedback

## 10 Further definition of regime shift - Define “abupt” and “Persistant”

## 11 Expand collaboration with Canadian Counterparts

## 12 Fall turnover date index

## 13 links between species availability inshore/offshore (estuarine conditions) and trends in recreational fishing effort?

## 14 Apex predator index (pinnipeds)

## 15 Forage availability index (Herring/Sandlance)

## 16 Trend analysis

## 17 Regime shifts in Social-Economic indicators

## 18 Linking Condition

## 19 Cumulative weather index

## 20 VAST and uncertainty

## 21 Seal index

## 22 Breakpoints

## 23 Management complexity

## 24 Shellfish growth/distribution linked to climate (system productivity)

## 25 Avg weight of diet components by feeding group

## 26 Mean stomach weight across feeding guilds

## 27 Recreational bycatch mortality as an indicator of regulatory waste

## 28 Sturgeon Bycatch

## 29 Decomposition of diversity drivers highlighting social components

## 30 Changing per capita seafood consumption as driver of revenue?

## 31 Nutrient input, Benthic Flux and POC(particulate organic carbon ) to inform benthic productivity by something other than surface indidcators

## 32 Relate OA to nutrient input; are there “dead zones” (hypoxia)?

## 33 indicators of chemical pollution in offshore waters?

## 34 how does phyto size comp affect EOF indicator, if at all?

## 35 Indicator of scallop pred pops poorly sampled by bts

## 36 Compare EOF (Link) thresholds to empirical thresholds (Large, Tam)

## 37 Time series analysis (Zooplankton/Forage fish) to tie into regime shifts

## 38 Optimum yeild for ecosystem

## 39 Re-evaluate EPUs

## 40 Incorporate social sciences survey from council

## 41 Biomass of spp not included in BTS

## 42 Reduce indicator dimensionality with multivariate statistics

## 43 Estuarine condition relative to power plants and temp

## 44 Young of Year index from multiple surveys

## 45 Inflection points for indicators

# References

