---
title:
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
output:
  pdf_document:
    includes:
      in_header: latex/header1.tex
    keep_tex: yes
bibliography: SOE2023.bib
csl: plos.csl
link-citations: yes
fontsize: 10pt
subparagraph: yes
urlcolor: blue
always_allow_html: true
---

```{r setup, include=FALSE}
#Default Rmd options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      dev = "cairo_pdf",
                      fig.align = 'center') #allows for inserting R code into captions

library(tidyverse)
library(readxl)
library(kableExtra)
library(ecodata)
library(googledrive)
library(pacman)
p_load(repmis, RCurl, DT, magrittr, gsubfn, stringr, tidyverse)

library(flextable)

set_flextable_defaults(
  font.size = 9,
  padding.bottom = 0,  
  padding.top = 1, 
  line_spacing = 1,
  font.family = "Palatino")


```

```{r, include = F}
#this is to wrap text around figures if we want
defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

<!--
NOTES

Use the code=readLines with the setup chunks from ecodata: need to do before each set of figs from a region/subgroup
-->

```{r, results='hide'}
 #  uncomment to get refreshed file from gdrive, otherwise it reads the one originally downloaded

 #  resultfile <- googledrive::drive_find(pattern = "SOE prioritization worksheet", type = "spreadsheet")
 # 
 # requests <- googledrive::drive_download(resultfile, type = "csv", overwrite = TRUE) %>% {read.csv(.$local_path)} 
 #   

 requests <- read.csv(here::here("SOE-response-memo/SOE prioritization worksheet.csv"))
```

# Introduction

In the table below we summarize all comments and requests with sources. The memo is now reorganized into categories of requests in descending order of overall Council priority. The Rank column summarizes priority and was derived from combined discussion with the Mid-Atlantic SSC ecosystem working group and a survey of selected MAFMC members coordinated by Council staff in July 2022, and updated based on discussion with the Mid-Atlantic SSC ecosystem and economic working groups and the New England SSC chair in July 2024. The Progress column briefly summarizes how we responded, with a more detailed response to each request in a section for each request category. In the Status column, "In SOE" indicates a change included in the report(s), and "In Catalog" indicates information included in the online indicator catalog.

```{r summtable, ft.arraystretch = 1}
# take from google drive to automatically update any changes
#googledrive::drive_download("Northeast IEA/State of the Ecosystem Reports/SOE 2020/Workshop & Meeting Notes/Request checklist", path="2019requests.xlsx", overwrite = TRUE)

#requests <- read_excel("2019requests.xlsx", sheet="MemoTable", range = 'A1:D31')

#requests$`Memo Section` <- paste0("[", requests$`Memo Section`, "](#",requests$`Memo Section`,")") #trying to make clickable links in the table to sections, not working
#adding format = "markdown" to kable statement made them links but they still didnt jump to report sections and messed up other formatting

requeststab <- requests %>%
  dplyr::select(Request, Year, Aggregate = Aggregate.Category, Project, Source, Status, Progress, Order) %>%
  dplyr::mutate(Source = str_trim(str_extract(Source, "[^-]+"))) %>%
  dplyr::group_by(factor(Aggregate, levels = c("System level thresholds/ref pts",
                                               "Management",
                                               "Short term forecasts",
                                               "Regime shifts",
                                               "Multiple system drivers",
                                               "Functional group level status/thresholds/ref pts",
                                               "Stock level indicators",
                                               "SOE admin"))
                  )%>% #, Project) %>%
  dplyr::mutate(#AvgOrder = mean(Order, na.rm=T),
                Rank = factor(case_when(Order > 0 & Order < 1.5 ~ "Highest", 
                                     Order >= 1.5 & Order < 1.75 ~ "High",
                                     Order >= 1.75 & Order < 2 ~ "Moderate",
                                     Order >= 2 & Order < 2.5 ~ "Low",
                                     Order >=2.5 ~ "Lowest",
                                     TRUE ~ "Unranked"),
                       levels=c("Highest", "High", "Moderate", "Unranked", "Low", "Lowest")))%>%
  dplyr::arrange(Rank, Order, .by_group = TRUE) %>% #AvgOrder, 
  # arrange(match(Status, c("In SOE", 
  #                         "In SOE-MAFMC",
  #                         "In SOE-MAFMC, In progress-NEFMC",
  #                         "In progress", 
  #                         "Not started")), desc(Year)) %>%
  dplyr::ungroup() %>%
  #dplyr::mutate("Memo Section" = rownames(.)) %>%
  #dplyr::mutate(Section = dplyr::group_indices(.)) %>%
  dplyr::select(Request, Year, Aggregate, Rank, Source, Status, Progress)

#requeststab$Section <- group_indices(requeststab)
  

# knitr::kable(requeststab, longtable = T, booktabs = TRUE, linesep = "") %>%
#   kable_styling(font_size=9, latex_options = c("hold_position", "striped", "repeat_header")) %>%
#   row_spec(0,bold=TRUE) %>%
#   column_spec(1, width="5cm") %>%
#   column_spec(2, width="1cm") %>%
#   column_spec(3, width="2cm") %>%
#   column_spec(4, width="5cm") %>%
#   column_spec(5, width="2cm")

flextable::as_grouped_data(requeststab, groups = "Aggregate") %>% 
  flextable::as_flextable(hide_grouplabel=TRUE) %>% 
  flextable::align(i = ~ !is.na(Aggregate), align = "left") %>% 
  flextable::bold(i = ~ !is.na(Aggregate), bold = TRUE) %>%
  flextable::theme_zebra() %>%
  flextable::bg(i = ~ !is.na(Aggregate), bg = "beige", part = "body") %>%
  flextable::hline(i = ~ !is.na(Aggregate)) %>%
  flextable::set_caption("State of the Ecosystem requests by category and Council priority.") %>%
  #flextable::autofit() 
  flextable::width(width = c(2.5,0.5,0.7,0.7,0.7,1.5))

```


# Overview of requests

This memo is organized into categories by topic, and categories are listed in descending order of overall priority based on approximate weighting within the category.  Therefore, a range of priority may be applied to individual requests within a category even though the entire category has an overall priority.  Priorities originally established in 2022 were reviewed by the SSCs in July 2024, and have been retained for 2025 in this memo. The request categories are:

#### System level thresholds/reference points
Includes requests to develop analytical methods that can be applied across all indicator types and operationalized for management advice. Much of this high priority methodological work is in progress.

#### Management
Includes analyses related to management performance. Work on this category is resource limited.

#### Short term forecasts 
Includes requests for biological and environmental forecasts. These forecasts include forthcoming CEFI products once they are tested. Work in this category started this year, but has been suspended because the person hired to do this work has been terminated.

#### Regime shifts 
Many analyses have been conducted and are in progress for individual ecosystem components, but a unifying framework with consistent methods is needed for the SOE.

#### Multiple system drivers
This category has the most requests. Most unranked requests from 2023-2024 are in this category. Further prioritization within this category is sorely needed.

#### Functional group level status/thresholds/reference points
Most of these requests are in progress. 

#### Stock level indicators
Requests for this information may be more appropriately directed to stock specific ecosystem products such as [Ecosystem and Socioeconomic Profiles (ESPs)](https://www.fisheries.noaa.gov/new-england-mid-atlantic/science-data/ecosystem-and-socioeconomic-profiles-northeast-united-states). 

New requests from 2023-2024 were either prioritized jointly through SSC discussion, or are listed without prioritization, while previously prioritized requests have been noted with adjustments to years in the table. We welcome further feedback on planned and continuing work.

## Adjustments to general SOE report structure

In response to feedback that the revision of the Climate and Ecosystem Risks section was useful, we have retained that section and made adjustments based on suggestions from both Council's staff in January 2025. The Climate and Ecosystem Risks section continues to center on management decisions. It includes 3 sections with modified titles to better align with management decisions: 
    +  Risks to managing spatially, highlighting distribution shifts in managed species with potential drivers
    +  Risks to managing seasonally, highlighting temporal shifts in managed species with potential drivers 
    +  Risks to setting catch limits, highlighting productivity and condition shifts, with potential drivers
    
In response to positive feedback on the 2023 Highlights section, we will continue to include a section summarizing conditions in the previous year. The 2024 Highlights section reviews new conditions, activities, and anomalous observations across the Northeast US from the past year. This section is summarized graphically on p.3 (Mid-Atlantic report) and p.4 (New England report).

We have retained and updated the [online indicator catalog](https://noaa-edab.github.io/catalog/index.html) provides a "deep dive" into each indicator, with multiple visualizations of the data and clearer links to the datasets in the [ecodata R package](https://noaa-edab.github.io/ecodata/) for increased transparency and ease of use by investigators throughout the region. We are also updating the [online technical documentation](https://noaa-edab.github.io/tech-doc/).  


```{r, results='asis', eval=FALSE}

for(j in 1:length(unique(requeststab$Aggregate))) {
  cat("##", unique(requeststab$Aggregate)[j], "\n")
  # for(i in 1:nrow(requeststab)) {
  #   #cat("##", requeststab$'Memo Section'[i], as.character(requeststab$Request[i]), "\n")
  #   
  #   cat("###", as.character(requeststab$Request[i]), "\n")
  #   cat("\n")
  # } 
  cat("\n")
}

```

## Summary of progress by category

### System level thresholds/reference points: highest priority

In July 2024, SSC review prioritized the following requests:  
    + maintain high priority on trend/threshold evaluation
    + express indicators relative to biological thresholds
    + standardize uncertainty language (IPCC)
    + longer term: simulation analysis of thresholds

#### Trend/threshold evaluation
The 2025 SOE addressed the first request by introducing a new short term trend analysis by Andy Beet (NEFSC) and applying it to all indicators where the analysis is appropriate. Short term trend analysis was applied the most recent 10 years of data for long (>30 year) time series, and to the full time series for indicators with fewer than 30 years of data for the 2025 SOE. The `ecodata` R package already incorporates long term trend estimation based on Hardison et al. @hardison_simulation_2019. This research found that trends were most robustly distinguished from autocorrelation in indicator time series of 30 years or longer. The new short term trend analysis addresses both autocorrelation in the short time series and missing data by applying a bootstrap analysis

Automated methods for evaluating time series inflection points, and identifying breakpoints (regimes) have been tested for inclusion in future SOEs. Kim Bastille (NEFSC) worked on methods to identify inflection points in indicator time series based on Large et al. @large_defining_2013 and @large_quantifying_2015. A standardized method has been implemented as a prototype and applied to several existing SOE indicators in 2022, but several questions on default approaches to be used across multiple indicators require more in depth analysis and review. A method for identifying breakpoints has been implemented by Kim Bastille and Laurel Smith (NEFSC) and a prototype analysis developed using SOE indicators in 2022. These methods are promising, but staff time has been devoted to other priorities for the past several years.

#### Biological thresholds 



#### Standard uncertainty language 
The SOE team has not had resources to review and incorporate standardized uncertainty language similar to that used by the IPCC for the 2025 report, although there is a desire to do so. One challenge is that not all contributed indicators include uncertainty estimates, so advice on treating a mix of indicators with various approaches to uncertainty would be most welcome. 

#### Indicator and threshold simulation analysis
Longer term simulation analysis of ecosystem thresholds is similarly of great interest to the SOE team, with a project currently in progress to evaluate ecosystem level overfishing (EOF) thresholds for the Northeast US region.  The EOF indicators were first presented in 2021 and were discussed in depth with the MAFMC SSC working group in April 2022 and February 2023. Considerable progress has been made on updating data inputs for the EOF indicators and planning for system level threshold analyses with the MAFMC SSC. After reviewing previous presentations of the EOF indicators, Andy Beet (NEFSC) reviewed solutions to several data input problems identified in July 2022 (menhaden landings were added and differences between different data sources were resolved). In 2023, estimates of regional productivity were added to calculate regional thresholds, for comparison with published global thresholds. An outstanding data input task is completing discard estimates for all species in the Northeast US, which is in progress. 

A simulation study is being planned to use the Northeast US Atlantis ecosystem model [@caracappa_northeast_2022] to investigate robustness of thresholds and determine how informative they can be. This portion of the research will likely address the MAFMC request to evaluate how phytoplankton size composition might affect the EOF indicator. It will also address SSC questions raised about tradeoffs between fishing for different species groups to address EOF, and how climate driven changes in transfer efficiency might be incorporated into or impact EOF indicators. In addition, the NEUS Atlantis model may be able to address the lower priority requests on nutrient input and benthic flux contributions to system productivity once model sensitivity analysis determines whether these model components behave reasonably. 


### Management: high priority

In July 2024, SSC review prioritized the following requests:  
    + include indicators for risk policy/risk assessment processes

The SOE team prioritized this request for 2025. The SOE already includes many indicators used in the MAFMC's EAFM risk assessment process, and several proposed for the newly approved NEFMC risk policy. Additional indicators associated with the MAFMC EAFM risk assessment have been added to the SOE ecodata package for 2025 (shoreside support indicators), and multiple new EAFM risk assessment indicators have also been included in 2025 (e.g., low trophic level prey abundance indices for benthic invertebrates and zooplankton that are combined with existing fish condition indicators to assess prey availability, social vulnerability indicators for all communities, and community climate risk indicators). Additional indicators are in development for offshore habitat risks based on stock-specific indicators, which will be presented in the MAFMC EAFM Risk Assessment document. These indicators will be added or adjusted after review by MAFMC for upcoming SOE and other ecosystem reporting products. 

Requests on management complexity and recreational bycatch mortality remain inadequately staffed to make progress. 

The request to re-evaluate Ecosystem Production Units (EPUs) was ranked lowest priority. We do not forsee having the resources to address this request, which is a large project, in the near future. 

### Short term forecasts: high priority

In July 2024, SSC review prioritized the following requests:  
    + include Climate Ecosystem Fisheries Initiative (CEFI) projections

The SSC working group ranked these new requests higher priority relative to Council members, but overall both ranked short term forecasting requests high priority. New resources to address this request came online at the NEFSC through the Climate Ecosystem Fisheries Initiative (CEFI) in 2024. This national effort seeks to link ocean model forecasts with products used in management. In the Northeast region, the SOE team was closely collaborating with CEFI modelers to test and present new products to the Councils and SSCs as they come online.

A preview of short term forecasts of bottom temperature for each EPU was initially reviewed in late February 2025. The intention was to continue to develop these forecasts and include retrospective forecasts alongside observations in the SOE Catalog to continue review for potential inclusion in the 2026 SOE. Due to CEFI staff loss, the timeline for this work is now significantly delayed. 

While using phytoplankton trends to forecast fish stocks may be feasibly simulation tested within the Atlantis modeling framework described above for EOF indicators, this relatively long term project would require dedicated effort to achieve. 

MAFMC has recently completed work on short term forecasts of species distributions for fisheries management are [in progress with Rutgers University and MAFMC](https://www.mafmc.org/s/Rutgers-Project-Overview.pdf), which may also address this request. Skill assessment of these forecasts, as well as determining the context in which they would be used (stock assessment projections? habitat projections? other uses?) would be needed to bring them into the management process (this is better developed for the ongoing Rutgers/MAFMC project). Incorporating short term forecasts into the SOE outside the ongoing Rutgers/MAFMC project would require a similar level of effort to the phytoplankton/fish forecasting project above. 

### Regime shifts: high priority

In July 2024, SSC review prioritized the following requests:  
   + characterize current conditions in context of expected short term change

Addressing the request to characterize expected short term change is linked to the short term forecasts category above. 

Adding information on regime shifts was considered a high priority by both the Council and SSC. Time series analysis of zooplankton to evaluate potential linked regime shifts is currently in progress, and multiple other regime shiftprojects may contribute to this. We are working to coordinate existing projects (see below) into a synthesis product for the SOE. Because the projects are on different timelines, it is difficult to give a target date for SOE synthesis. However, a workshop is scheduled for May 2024 to review and synthesize methods and results for the Northeast Region. The workshop report was not available for the 2025 SOE. With these publications complete, some synthesis may be presented in the following SOE cycle. 

Regime shifts in socio-economic indicators may be addressed in the ongoing work described above by John Walden and Geret DePiper (NEFSC) integrating multiple indicators into Index Numbers. Once the structure of the Index Numbers is determined, these time series can be evaluated for change points using any of the methods described in the table above. 

### Multiple system drivers: moderate-high priority

In July 2024, SSC review prioritized the following requests:  
    + profits vs revenue: provide incomplete net revenue and index of costs
    + clarify objectives and terminology for fishing community engagement and social indicators
        + time series of community indicators
        + social and economic linkages to climate
        + consider appropriate scale for indicators

This category contains a wide array of requests with many projects currently in progress. There were two requests ranked high priority, five ranked moderate priority, nine unranked because they are newer requests, and eight ranked low or lowest priority. Given the number of SOE requests, those ranked lowest priority that have not already been started are unlikely to be addressed. 

An unranked request to review direct indicators of net vs. gross revenue indicators is in progress. We can calculate net revenue for ~ 1/2 of the revenue generated within the Greater Atlantic Region. The trends between the total gross revenue and gross revenue for which we can estimate costs are different. In addition, the net revenue looks to be just a scaled gross revenue metric, with trends staying the same, just the magnitude changing (Fig. \ref{fig:profit}). 

```{r profit, fig.cap="Net revenue (left) and cost coverage (right) in the Greater Atlantic Region.", out.width = "49%", fig.show='hold'}

#knitr::include_graphics(here::here("SOE-response-memo/images/Net_revenue.pdf"))
#knitr::include_graphics(here::here("SOE-response-memo/images/Cost_Coverage.pdf"))
```

Another high priority request in this category is incorporating the ongoing fish condition project and associated analyses into the SOE. Regime shift analyses of fish condition may be available for the 2025 SOE. Initial work linking multiple SOE low trophic time series with fish condition using dynamic structural equation modeling [@thorson_dynamic_2024] was started in February 2024. With additional resources, this approach may be presented to synthesize drivers of fish condition across multiple species in a future SOE.  



### Functional group level status/thresholds/ref pts: moderate priority

In July 2024, SSC review prioritized the following requests:  
    + not specifically prioritized
    + include more aggregations for biomass and landings (Council-managed, status)


Requests in this category were considered moderate to low priority by the SSC and Council. However, many were already in progress prior to ranking, and one has been included in the 2023 SOE.

The NEFMC requested a forage availability index (including both managed species such as herring and unmanaged species such as sandlance). A spatial index of forage availability was developed for the bluefish research track assessment as described above. This index was partitioned into EPUs and presented in the 2023 and 2024 SOEs. An index of forage center of gravity was also included as a potential driver of distribution shifts in the 2024 Risks to Spatial Management section.

Gray seal pup count indices are already included in the NEFMC SOE, and indices of populations for other seals and apex predators are in development by the protected species branch. These additional indices were not ready for the 2024 report. 

Investigating time series of biomass for species not well represented in bottom trawl surveys was partially addressed by the forage index included in the 2023-2024 reports. However, only a subset of forage species are not well represented in bottom trawl surveys, and other species that are not forage are also not well represented in bottom trawl surveys. This request was ranked lowest priority by the Council and SSC, and given the difficulty of synthesizing data on poorly sampled species, is unlikely to be addressed in the near future.

### Stock level indicators: moderate priority

In July 2024, SSC review prioritized the following requests:  
    + not specifically prioritized 
    + cross reference ESP products where appropriate


Requests in this category were ranked moderate to lowest priority by the SSC and Council. Indicators of this nature would be well suited to Ecosystem and Socioeconomic Profiles (ESP) developed during research track assessments for individual stocks. Some aspects of these indicators may benefit SOE reporting as well.

One request, linking shellfish growth and distribution to climate change and system productivity, is in progress. Alexis Hollander (VIMS) completed her thesis on surfclam growth in relation to bottom temperature in 2022, and information from this work can likely be included in the 2025 SOE, pending publication of student thesis results. 

The request for indicators of scallop predators that are poorly sampled by bottom trawls is similar to the request in the category above addressing all species not well sampled by bottom trawls. It is possible that this request could be clarified and addressed during a scallop research track assessment.

A new unranked request to evaluate the impacts of climate change on survey catchability is being addressed in some research track stock assessments. 

The request for a sturgeon bycatch indicator was ranked lowest priority by the SSC and Council, so is unlikely to be addressed in the near future.

### SOE admin: unranked priority

These requests were not ranked; however, many are in progress.

In 2024, both SSCs asked for criteria used to include information in the new "Highlights" sections. As noted in the [2024 Highlights catalog page](https://noaa-edab.github.io/catalog/observation_synthesis_2024.html), Observations were solicited from contributors to the State of the Ecosystem reports, colleagues at the Northeast Fisheries Science Center, academic and management partners, and the fishing industry. In late 2024 an email account was set up so that anyone can provide observations to the SOE team: <northeast.ecosystem.highlights@noaa.gov>. Items selected for inclusion in the synthesis include observations that were at or near their respective time series record (high or low) or had a large departure from the recent trends, were reported by multiple contributors, affected fishing efforts, or were considered newsworthy. 

Investigation of uses of the SOE as requested by the MAFMC SSC is in progress with the assistance of NOAA communications experts using a combination of website analytics and citation information. This information has been compiled as part of an in progress manuscript reviewing use of ecosystem reports across NMFS. We hope to have an update on uses of the Northeast SOEs for the 2025 report/request memo. 

The restructuring of this memo according to prioritization is intended to partially address the requests for timelines on in progress SOE requests by the NEFMC SSC. While not all project timelines are currently available, we have reported estimates in this document where possible. In addition, the effort to prioritize requests in 2022 ensures that limited resources are applied to the highest priority issues.


# References

